Using numerical modelling to investigate temperature and wave exposure as drivers of macroalgal characteristics

### 2019-08-21 morphology_paper.Rmd
literature review added. updates to rmd script added. issues with pushing to repro rectified.

```{r map study sites, echo=FALSE, message=FALSE, warning=FALSE}
library(png)
library(grid)

@@

```{r schematic, echo=FALSE, fig.cap='Schematic of *L. pallida* and *E. maxima* morphology.'}

@@

#temperature and lamanaria  
LT1 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaWeight, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
  ylab("Lamina weight") +
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT2 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaLength, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
  ylab("Lamina length") +
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT3 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaThickness, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
  ylab("Lamina thickness") +
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT4 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeMass, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT5 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeLength, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT6 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeDiameter, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
  ylab("Stipe diameter (cm)") +
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT7 <- ggplot(data = lam_data, aes(x = temp_regime, y = digits, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
  ylab("No. of digits") +
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT8 <- ggplot(data = lam_data, aes(x = temp_regime, y = thallusMass, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
  ylab("Thallus mass (g)") +
  scale_fill_discrete(name="Temperature regime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())
LT9 <- ggplot(data = lam_data, aes(x = temp_regime, y = totalLength, fill = temp_regime)) +
  stat_compare_means(method = "wilcox.test") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank())

@@

LW1 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Lamina weight (g)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level"))

LW2 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Lamina length (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level"))

LW3 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Lamina thickness") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level"))

LW4 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Stipe mass (g)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level"))

LW5 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeLength, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Stipe length (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level"))

LW6 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Stipe diameter (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level"))

LW7 <- ggplot(data = lam_data, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("No. of digits") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level"))

LW8 <- ggplot(data = lam_data, aes(x = wave_exposure, y = thallusMass, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Thallus mass (kg)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level"))

LW9 <- ggplot(data = lam_data, aes(x = wave_exposure, y = totalLength, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  ylab("Total length (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) 
```

```{r laminaria temp plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by temperature regime.', fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
```

```{r lamanaria wave plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by wave exposure category.', fig.height=6, fig.width=6, warning=FALSE}
ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
```

@@

  stat_compare_means(method = "wilcox.test") + 

@@

  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) 

@@

  stat_compare_means(method = "wilcox.test") + 

@@

  stat_compare_means(method = "kruskal.test") + 

@@

  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +

@@

  stat_compare_means(method = "kruskal.test") + 

@@

  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +

@@

```{r deep ecklonia temp plot, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E. maxima* grouped by temperature.', fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
```

```{r deep ecklonia wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E. maxima* grouped by wave exposure.', fig.height=6, fig.width=6, warning=FALSE}
ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom")
```

@@

  stat_compare_means(method = "wilcox.test") + 

@@

@@

```{r shallow ecklonia temp plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by temperature regime.', fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = FALSE, legend = "bottom")
```

```{r shallow rda wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by wave exposure.', fig.height=6, fig.width=6, warning=FALSE}
ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = TRUE, legend = "bottom")
```

@@

# Plot RDA
rda_eck_deep <- ggord(rda_eck, eck_meta_rda$region ,ellipse = TRUE, hull = FALSE, ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.95)  # looking at the raw code, this is plotting the 'wa scores', the blue 
```

@@

# Plot RDA
rda_eck_shallow <- ggord(rda_eck_shallow,meta_shallow_rda$region,ellipse = TRUE, hull = FALSE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9, coord_fix = TRUE)  # looking at the raw code, this is plotting the 'wa scores', the blue

@@

```{r rda plots, echo=FALSE, fig.height=9, fig.width=9, message=FALSE, warning=FALSE}
ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, common.legend = TRUE, ncol = 3, nrow = 1, legend = c("bottom"))
```

@@

This study considered important drivers of kelp morphology that have been identified in previous research (Fowler-Walker et al. 2006; Miller et al. 2011; Fowler-Walker et al. 2005; Stewart et al. 2009; Wing et al. 2007; Molloy & Bolton 1996; Pedersen & Nejrup 2012; Wernberg & Thomsen 2005), namely temperature and wave exposure, around the Cape Peninsula on two prolific kelp species. Unlike previous research, this study used a numerical model to provide the various variables of wave exposure. We also considered wind in addition to temperature and wave variables, since wind has been shown to influence the extension capabilities of canopy kelp species (Denny et al. 1997). Furthermore, the protrusion of the primary blade of *E. maxima* out of the water may result in some direct effects of wind-drag on morphology. The results show that wave exposure the main driver of kelp morphology with temperature and wind playing only minor roles. Furthermore, the investigation of differences in morphological characteristics between shallow and deep populations of *E. maxima *suggests that in low wave exposure environments the role of temperature as a morphological driver increases.

@@

Past research shows that macroalgal morphological characteristics are largely driven by a wave exposure while others have suggested a complex interaction between various environmental variables. We confirm that wave variables are the main driver of kelp morphological characteristics in both *E. maxima *and *L. pallida* populations around the Cape Peninsula, and differences in shallow and deep populations of *E. maxima* may reflect small differences in wave and temperature variables as drivers of kelp morphological characteristics.

>>>>

```{r map study sites, echo=FALSE, fig.height=7, fig.width=6, message=FALSE, warning=FALSE}
library(png)
library(grid)

@@

```{r schematic, echo=FALSE, fig.cap='Schematic of *L. pallida* and *E. maxima* morphology.', fig.height=7, fig.width=7}

@@

#temperature and lamanaria  
LT1 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaWeight, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina weight") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LT2 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaLength, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina length") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LT3 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaThickness, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina thickness") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LT4 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeMass, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Stipe mass (g)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LT5 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeLength, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Stipe length (cm)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LT6 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeDiameter, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Stipe diameter (cm)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LT7 <- ggplot(data = lam_data, aes(x = temp_regime, y = digits, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
  xlab("") + 
  ylab("No. of digits") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LT8 <- ggplot(data = lam_data, aes(x = temp_regime, y = thallusMass, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Thallus mass (g)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LT9 <- ggplot(data = lam_data, aes(x = temp_regime, y = totalLength, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Total length (cm)") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

@@

LW1 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
  ylab("Lamina weight (g)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW2 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
  ylab("Lamina length (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW3 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
  ylab("Lamina thickness") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW4 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
  ylab("Stipe mass (g)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW5 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeLength, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
  ylab("Stipe length (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW6 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
  ylab("Stipe diameter (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW7 <- ggplot(data = lam_data, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
  ylab("No. of digits") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW8 <- ggplot(data = lam_data, aes(x = wave_exposure, y = thallusMass, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
  ylab("Thallus mass (kg)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")
LW9 <- ggplot(data = lam_data, aes(x = wave_exposure, y = totalLength, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
  ylab("Total length (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")
```

```{r laminaria temp plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by temperature regime.', fig.height=7, fig.width=8, message=FALSE, warning=FALSE}
ggarrange(LT1, LT2,LT3,LT4,LT5,LT6,LT7,LT8,LT9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
```

```{r lamanaria wave plots, echo=FALSE, fig.cap='Boxplots of *L. pallida* morphological characteristics group by wave exposure category.', fig.height=7, fig.width=8, warning=FALSE}
ggarrange(LW1, LW2,LW3,LW4,LW5,LW6,LW7,LW8,LW9, labels = c("AUTO"), common.legend = TRUE, legend = "bottom")
```

@@

  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 

@@

  theme_bw() +
  theme(axis.text.x = element_blank()) + 

@@

  stat_compare_means(method = "wilcox.test", label = "p.format") + 

@@

  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 

@@

  theme_bw() +
  theme(axis.text.x = element_blank()) + 

@@

  stat_compare_means(method = "kruskal.test", label = "p.format", label.y.npc = c('top')) + 
  theme_bw() +
  theme(axis.text.x = element_blank()) + 

@@

@@

```{r deep ecklonia temp plot, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E. maxima* grouped by temperature.', fig.height=8, fig.width=8, message=FALSE, warning=FALSE}
ggarrange(ET1,ET2,ET3,ET4,ET5,ET6,ET7,ET8,ET9,ET10, labels = "AUTO", legend = FALSE, ncol = 3, nrow = 4)
```

```{r deep ecklonia wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for deep population *E. maxima* grouped by wave exposure.', fig.height=8, fig.width=8, warning=FALSE}
ggarrange(EW1,EW2,EW3,EW4,EW5,EW6,EW7,EW8,EW9,EW10, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3, nrow = 4)
```

@@

  stat_compare_means(method = "wilcox.test", label = "p.format", label.y.npc = c('top')) + 
  theme_bw() +
  theme(axis.text.x = element_blank()) + 

@@

  theme_bw() +
  theme(axis.text.x = element_blank()) + 

@@

```{r shallow ecklonia temp plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by temperature regime.', fig.height=7, fig.width=8, message=FALSE, warning=FALSE}
ggarrange(EST1,EST2,EST3,EST4,EST5,EST6,EST7,EST8,EST9,EST10, labels = "AUTO", common.legend = FALSE, legend = "bottom", ncol = 3, nrow = 4)
```

```{r shallow rda wave plots, echo=FALSE, fig.cap='Boxplot of morphological characteristics for shallow population *E. maxima* grouped by wave exposure.', fig.height=7, fig.width=8, warning=FALSE}
ggarrange(ESW1,ESW2,ESW3,ESW4,ESW5,ESW6,ESW7,ESW8, ESW9, labels = "AUTO", common.legend = TRUE, legend = "bottom", ncol = 3,nrow = 3)
```

@@

# Plot RDA
rda_eck_deep <- ggord(rda_eck, eck_meta_rda$region ,ellipse = TRUE, hull = FALSE, ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.95)  
```

@@

# Plot RDA
rda_eck_shallow <- ggord(rda_eck_shallow,meta_shallow_rda$region,ellipse = TRUE, hull = FALSE, grp_title = "Sites", ptslab = TRUE, parse = TRUE, addsize = 3.5, ext = 0.9, coord_fix = TRUE)

@@

```{r rda plots, echo=FALSE, fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
ggarrange(lam_rda, rda_eck_deep, rda_eck_shallow, ncol = 1, nrow = 3, label = "AUTO")
lam_rda
rda_eck_deep
rda_eck_shallow
```

@@

This study considered important drivers of kelp morphology that have been identified in previous research (Fowler-Walker et al. 2006; Miller et al. 2011; Fowler-Walker et al. 2005; Stewart et al. 2009; Wing et al. 2007; Molloy & Bolton 1996; Pedersen & Nejrup 2012; Wernberg & Thomsen 2005), namely temperature and wave exposure, around the Cape Peninsula on two prolific kelp species. Unlike previous research, this study used a numerical model to provide the various variables of wave exposure. We also considered wind in addition to temperature and wave variables, since wind is an important component of surface driven waves. The results show that specific variables of wave exposure are the main drivers of kelp morphology with temperature and wind playing only minor roles. Furthermore, the investigation of differences in morphological characteristics between shallow and deep populations of *E. maxima * suggests that in low wave exposure environments the role of temperature as a morphological driver increases, such as in False Bay.

@@

Past research shows that macroalgal morphological characteristics are largely driven by a wave exposure while others have suggested a complex interaction between various environmental variables. We confirm that wave variables are the main driver of kelp morphological characteristics in both *E. maxima *and *L. pallida* populations around the Cape Peninsula, and differences in shallow and deep populations of *E. maxima* may reflect small differences in wave and temperature variables as drivers of kelp morphological characteristics. The response of kelp morphological characteristics to wave exposure was evident and both species show a tactics based on the magnitde of wave exposure. Strength increasing traits were exibited for lower exposure levels while a 'go-with-the-flow' tactic for moderate levels of wave exposure. Denny and Galord ??? showed that increased flexiblity or a 'go-with-the-flow' tactic has a threshold value, where increased stipe length actaully increase drag under high wave exposure scenarios. Therefore, under high wave exposure scenarios size reducing traits, such as reduced surface area of the overall plant should be favoured instead. This was evident in the results which showed that stipe length decreases under high wave exposure categories. This suggests that how morphological characteristcs manifest themselves is dependent on the magnitude of wave exposure. Furthermore, the results suggest that in the absence of wave exposure, temperatre becomes the dominant morphological driver. This is evident from the redundancy analysis for each species and deep populations of *E. maxima* which show clustering of study sites according to the region; wave variables drive morphological characteristics for kelps from the western region and temperature variables for kelps from the eastern region. Clustering of sites according to region overlapped almost completey for the RDA analysis of shallow *E. maxima* suggesting that for shallow kelp the effect of wave exposure is less variable. Temperature variables were not a significant driver of kelp morphology in this study. Although temperature plays an important role in distribution (Bolton et al. 2012; Miller et al. 2011; Rinde et al. 2014) of kelp and physiological functioning of adults and gametophytes (Mohring et al. 2014; Steneck et al. 2002; Gerard 1997; Bearham et al. 2013; Smale & Moore 2017), there is little evidence that temperature is an important driver of morphological variation. However, we suggest that temperature plays a larger role in determining kelp morphology in areas where wave energy does not exert large amounts of impact.
